School Socioeconomic Status Context and Social Skills in Children

### 2020-10-23 manuscript.tex
Full results section

After decades of research and intervention aimed at reducing educational and occupational inequality, inequalities persist (Reardon, 2011). In response to this bleak picture, there has been a flurry of interest in social-emotional competencies driven by the research of Nobel prize-winning economist James Heckman. (Heckman, 2006) argues that research and intervention on reducing educational inequality have focused too narrowly on ``cognitive'' abilities. Yet cognitive abilities are not the only pathway through which low socioeconomic status may stifle educational and occupational attainment. For example, social skills are a potentially powerful explanation for socioeconomic status gaps in educational and occupational attainment (Gutman \& Schoon, 2013). Children from low socioeconomic backgrounds (here after low SES children\footnote{We acknowledge that children themselves do not have a socioeconomic status. Rather they are raised in an environment which is shaped by their parents' socioeconomic status (SES). However, for brevity we refer to children as low SES or high SES hereafter. Likewise, in place of low school average SES and high school average SES we refer to disadvantaged and advantaged schools respectively.}) appear to enter school with poorer social skills (e.g., Jerrim \& Sims, 2019). This is concerning as childhood social skills predict a wide variety of later life outcomes (e.g., Jones, Greenberg, \& Crowley, 2015). Inspired by identity economics, our research asks whether the school context in the stratified school system of Australia may exacerbate this issue. In particular, we explore whether school average socioeconomic status predicts social skills in Year 3 (age 8)---using measures of peer problems, conduct problems, and prosocial behavior. We consider if this association is independent of children's incoming social skills at age 4, children's own socioeconomic status, and a set of demographic and academic covariates in a representative longitudinal sample of Australian children.

@@

Of the skills that employers are looking for when hiring candidates, social skills are some of the most desirable (Rios, Ling, Pugh, Becker, \& Bacall, 2020). Social skills also seem to be a viable target for intervention with meta-analysis showing social and emotional learning programs have moderate effects on improving key social skills like reducing conduct problems and improving social-emotional learning skills and prosocial behavior (Durlak, Weissberg, Dymnicki, Taylor, \& Schellinger, 2011). Social skills appear to be worthwhile targets from a societal perspective given their relationship to academic achievement (Corcoran, Cheung, Kim, \& Xie, 2018) and their role in predicting adult employment, crime, public assistance, substance abuse, and mental health (Jones et al., 2015). Among young children, it is increasingly common to focus on social skills as measured by the Strengths and Difficulties Questionnaire (see Datta Gupta \& Simonsen, 2010; Gutman \& Schoon, 2013; Jerrim \& Sims, 2019). We use this questionnaire to measure peer problems, conduct problems, and prosocial behavior. We model social skills using both parent and teacher reports.

@@

(Jerrim \& Sims, 2019) provide one test of the potential assimilative effect of schools on social skills. They found that living in an area with selective schools is associated with better social skills when compared to children living in districts without a selective school. They also found some assimilative effects of school average SES predicting social skills using the total score of the strengths and difficulties questionnaire (SDQ). The assimilation effects they found, however, were weak. This may be because the authors used the SDQ total score---which includes a range of social and emotional behaviors---rather than focusing on specific social skill components of the SDQ. Further, assimilative effects may be non-linear. Put another way, the assimilative power of school contexts may not be evenly distributed across the socioeconomic distribution. This is an important consideration that research has not hitherto considered, though differential assimilation across the socioeconomic status gradient has been theorized (Gradstein \& Justman, 2005).

@@

\begin{figure}
\centering
\includegraphics{/Users/phparker/Dropbox/Projects_Research/CURRENT_WORKING/social_skills_2020/fig/flow.pdf}
\caption{\label{fig:consort}Participant flow diagramme}
\end{figure}

We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC). LSAC is a government-run study of a representative sample of Australian children between who were zero-one (B-cohort) and four-five (K-Cohort) years of age in 2004. Both cohorts of children have been followed every two years since (AIFS, 2015). LSAC includes linked administration records for student performance in a standardized national numeracy test. We also use government collected data on school average socioeconomic status. Many children in the sample were the only child surveyed (25.75\%) or were one of two children surveyed (20.96\%) in their school. By using administrative records, we gain school average socioeconomic status based on the child's \(\text{\underline{whole}}\) school grade from a high-quality source, thus avoiding sampling bias present in most school context research. We excluded participants who were not in school (e.g., were home schooled, or were otherwise missing an identifier of which school they attended), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability). Together our total sample was 5440 (52.08\% boys) children in Year 3 at school (\textasciitilde age 8). In the vast majority of cases (98\%), parent data came from mothers. A participant flow diagram is present end in Figure~\ref{fig:consort} including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).

@@

Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ (\emph{Strengths and difficulties questionnaire}, 1997). We explored social skills at age 4 (for prior social skills) and age 8 (as primary outcomes) as reported by the child's \emph{parent} and the child's \emph{teacher}. The SDQ asks respondents to rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true). Questions ask about the child's peer problems (``Rather solitary, tends to play alone''), conduct problems (``Often fights with other children or bullies them''), and prosociality (``Is kind to younger children''). We used the total scores for these scales as developed by the LSAC administrators. Scores ranged from 0 to 10. These scores were heavily left censored (see Figures 2-3) for peer and conduct problems with a preponderance of students being scored as a zero by their parent or teacher. prosocial behaviors were heavily right censored. Parents and teachers had moderate agreement for children at age 8 with correlations ranging from r = .40 for peer and r = .39 for conduct problems to r = .28 for prosocial behavior.

The SDQ items measure a collection of associated behaviours and as such we treat them as index or a formative variable. In this case common reliability metrics do not make much sense; nor do reflective models like confirmatory factor analysis. Instead, we would want items that are not too redundant with each other. For formative scales a variable inflation factor (VIF) of less than 10 has been suggested has a cut-off score; with VIF values of less than 3.3 showing an item makes an excellent contribution to the index (Diamantopoulos, Riefler, \& Roth, 2008; Thongrattana, 2010). In our case no VIF for any item for any social skills construct was higher than 2. Likewise, a limit of 30 on the condition index has been suggested as a good measure of the reliability of a formative construct (Thongrattana, 2010). This was the case for all our social skills constructs at both age 4 and age 8 and for both parent and teacher report. Parallel analysis suggested that all social skills constructs could be represented by a single principal component (but in all cases more than one reflective factor). The only exception was age 4 peer problems were parallel analysis suggested two components but the eigenvalue on the second component was barely greater than the simulated data suggesting a single index was likely still appropriate.

@@

Numeracy for the child and for their complete school grade was taken from administration records of National Assessment Program -- Literacy and Numeracy (NAPLAN) test results. The relevant NAPLAN tests for this study are given to all eligible children in the country in Year 3 (age 8). The tests are scaled so they are comparable across age cohort and across year grade. These high-stakes tests have a mean of 500 and a standard deviation of 100. Included in the analysis was the first principal component of the numeracy, grammar, spelling, reading, and writing scores from NAPLAN\footnote{Parallel analysis suggested a single principal component was sufficient to account for the variation in the NAPLAN scores.}. For verbal ability we used the Peabody Picture Vocabulary test (\emph{Peabody picture vocabulary test--fourth edition}, 2007) given to participants at age 4 (i.e., at or just before school commencement). Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender---all measured at age 4. The sector (government or non-government) of the school the child attended was also included as a covariate.

@@

There was a relatively small amount of missing data for most variables. However, given the data were longitudinal and included both linked administrative records and data collected from both the child's teacher and parent, missing data was inevitable. Attrition was offset by using applying combined sample and attrition weights constructed for the age 8 sample. Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14\%. All other variables had 6\% missing data or less. To account for this, we constructed five imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package (Honaker, King, \& Blackwell, 2011). These imputations were used in all analyses. All analyses were done using Bayes via the BRMS package in R (Bürkner, 2017).

All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobit regression. Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails ((see Kleiber \& Zeileis, 2008). As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models. The multilevel nature of the data (children nested within schools) was addressed by the inclusion of a random intercept for schools.

@@

In order to provide a context for the main analyses we first aimed to show a) that richer children enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools. Our findings show that high SES children do enter school with better social skills. Student SES is correlated with less peer (B = -.284 {[}-.345, -.222{]}) and conduct (B = -.371 {[}-.438, -.304{]}) problems. They also had slightly higher prosocial behavior (B = .074 {[}.009, .139 {]}). Further, student SES was strongly related to the school average SES of the school they attended (r = .480 {[}.458, .499{]}; see Figure 1).

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{llrrr}
\caption*{
\large Table 1\\ 
\small Effect of School Average SES on Social Skills\\ 
} \\ 
\toprule
Outcome & Report Source & Estimate & -95\% CI & +95\% CI \\ 
\midrule
conduct & teacher & $-0.153$ & $-0.259$ & $-0.043$ \\ 
social & teacher & $0.148$ & $0.053$ & $0.243$ \\ 
peer & teacher & $-0.116$ & $-0.204$ & $-0.027$ \\ 
conduct & parent & $-0.108$ & $-0.172$ & $-0.045$ \\ 
social & parent & $0.087$ & $0.015$ & $0.157$ \\ 
peer & parent & $-0.134$ & $-0.206$ & $-0.063$ \\ 
\bottomrule
\end{longtable}

@@

We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4. Results are presented under in Table \ref{tab:table-1}. Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3). School average SES negatively predicted conduct problems and positively predicted prosocial behavior for both parent and teacher reported social skills. School average SES also negatively predicted peer problems but only for parent reports.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{llrrr}
\caption*{
\large Table 2\\ 
\small Effect of School Average SES by Student SES on Social Skills\\ 
} \\ 
\toprule
Outcome & Report Source & Estimate & -95\% CI & +95\% CI \\ 
\midrule
conduct & teacher & $-0.160$ & $-0.269$ & $-0.049$ \\ 
conduct & teacher & $0.151$ & $0.063$ & $0.239$ \\ 
social & teacher & $0.156$ & $0.060$ & $0.250$ \\ 
social & teacher & $-0.094$ & $-0.170$ & $-0.018$ \\ 
peer & teacher & $-0.125$ & $-0.214$ & $-0.036$ \\ 
peer & teacher & $0.116$ & $0.041$ & $0.189$ \\ 
conduct & parent & $-0.112$ & $-0.174$ & $-0.049$ \\ 
conduct & parent & $0.059$ & $0.006$ & $0.112$ \\ 
social & parent & $0.090$ & $0.020$ & $0.159$ \\ 
social & parent & $-0.031$ & $-0.088$ & $0.026$ \\ 
peer & parent & $-0.143$ & $-0.214$ & $-0.071$ \\ 
peer & parent & $0.118$ & $0.058$ & $0.177$ \\ 
\bottomrule
\end{longtable}

@@

The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table 2). Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior. Interactions are plotted in Figures 2-4. Likewise, counterfactual expected score differences on the latent continuous social skill (underlying the censored social skill in the Tobit model) are given in Table 4. These results show that low SES children appear to be most influenced by school average SES. Middle to high SES children appeared to have largely similar scores regardless of the school they were in (see Figures 2-4). For example, both parents and teachers report that low SES children had approximately half a point more peer problems in more SES disadvantaged schools than in more SES advantaged schools (see Table 4). High SES children had almost no difference in peer problems regardless of the school they were in.

@@

\leavevmode\hypertarget{ref-diamantopoulos2008}{}%
Diamantopoulos, A., Riefler, P., \& Roth, K. (2008). Advancing formative measurement models. \emph{Journal of Business Research}, \emph{61}(12), 1203--1218. \url{https://doi.org/10.1016/j.jbusres.2008.01.009}

@@

@@

@@

\leavevmode\hypertarget{ref-thongrattana2010}{}%
Thongrattana, P. (2010). Assessing reliability and validity of a measurement instrument for studying uncertain factors in thai rice supply chain. In. Presented at the SBS hdr student conference. Retrieved from \url{https://ro.uow.edu.au/sbshdr/2010/papers/4/}

>>>>

After decades of research and intervention aimed at reducing educational and occupational inequality, inequalities persist (Reardon, 2011). In response to this bleak picture, there has been a flurry of interest in social-emotional competencies driven by the research of Nobel prize-winning economist James Heckman. (Heckman, 2006) argues that research and intervention on reducing educational inequality have focused too narrowly on ``cognitive'' abilities. Yet cognitive abilities are not the only pathway through which low socioeconomic status may stifle educational and occupational attainment. For example, social skills are a potentially powerful explanation for socioeconomic status gaps in educational and occupational attainment (Gutman \& Schoon, 2013). Children from low socioeconomic backgrounds (here after low SES children\footnote{We acknowledge that children themselves do not have a socioeconomic status. Rather they are raised in an environment which is shaped by their parents' socioeconomic status (SES). However, for brevity we refer to children as low SES or high SES hereafter. Likewise, in place of low school average SES and high school average SES we refer to disadvantaged and advantaged schools respectively.}) appear to enter school with poorer social skills (e.g., Jerrim \& Sims, 2019). This is concerning as childhood social skills predict a wide variety of later life outcomes (e.g., Jones, Greenberg, \& Crowley, 2015). Inspired by identity economics, our research asks whether the school context in the stratified school system of Australia may exacerbate this issue. In particular, we explore whether school average socioeconomic status predicts social skills in Year 3 (age 8)---using measures of peer problems, conduct problems, and prosocial behavior. Using prospective representative data, integrated with government administrative data, we consider if this association is independent of children's incoming social skills at age 4, children's own socioeconomic status, and a set of demographic and academic covariates.

@@

Of the skills that employers are looking for when hiring candidates, social skills are some of the most desirable (Rios, Ling, Pugh, Becker, \& Bacall, 2020). Social skills also seem to be a viable target for intervention with meta-analysis showing social and emotional learning programs have moderate effects on improving key social skills like reducing conduct problems and improving social-emotional learning skills and prosocial behavior (Durlak, Weissberg, Dymnicki, Taylor, \& Schellinger, 2011). Social skills appear to be worthwhile targets from a societal perspective given their relationship to academic achievement (Corcoran, Cheung, Kim, \& Xie, 2018) and their role in predicting adult employment, crime, public assistance, substance abuse, and mental health (Jones et al., 2015). Among young children, it is increasingly common to focus on social skills as measured by the Strengths and Difficulties Questionnaire (see Datta Gupta \& Simonsen, 2010; Gutman \& Schoon, 2013; Jerrim \& Sims, 2019). We use this questionnaire to measure peer problems, conduct problems, and prosocial behavior. We model social skills using both parent and teacher reports. Measures of SES at the student level are taken from in-person interviews with parents and school SES (as well as student achievement) is taken from government administrative records.

@@

Jerrim and colleagues (2019) provide one test of the potential assimilative effect of schools on social skills. They found that living in an area with selective schools is associated with better social skills when compared to children living in districts without a selective school. They also found some assimilative effects of school average SES predicting social skills using the total score of the strengths and difficulties questionnaire (SDQ). The assimilation effects they found, however, were weak. This may be because the authors used the SDQ total score---which includes a range of social and emotional behaviors---rather than focusing on specific social skill components of the SDQ. Further, assimilative effects may be non-linear. Put another way, the assimilative power of school contexts may not be evenly distributed across the socioeconomic distribution. This is an important consideration that research has not hitherto considered, though differential assimilation across the socioeconomic status gradient has been theorized (Gradstein \& Justman, 2005).

@@

\begin{figure}
\centering
\includegraphics{/Users/philipparker/Dropbox/Projects_Research/CURRENT_WORKING/social_skills_2020/fig/flow.pdf}
\caption{\label{fig:consort}Participant flow diagramme}
\end{figure}

We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC). LSAC is a government-run study of a representative sample of Australian children between who were zero-one (B-cohort) and four-five (K-Cohort) years of age in 2004. Both cohorts of children have been followed every two years since (AIFS, 2015). LSAC includes linked administration records for student performance in a standardized national numeracy test. We also use government collected data on school average socioeconomic status. Many children in the sample were the only child surveyed (25.75\%) or were one of two children surveyed (20.96\%) in their school. By using administrative records, we gain school average socioeconomic status based on the child's \(\text{\underline{whole}}\) school grade from a high-quality source, thus avoiding sampling bias present in much school context research. We excluded participants who were not in school (e.g., were home schooled, or were otherwise missing an identifier of which school they attended), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability). Together our total sample was 5440 (52.08\% boys) children in Year 3 at school (\textasciitilde age 8). In the vast majority of cases (98\%), parent data came from mothers. A participant flow diagram is present end in Figure~\ref{fig:consort} including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).

@@

Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ (\emph{Strengths and difficulties questionnaire}, 1997). We explored social skills at age 4 (for prior social skills) and age 8 (as primary outcomes) as reported by the child's \emph{parent} and the child's \emph{teacher}. The SDQ asks respondents to rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true). Questions ask about the child's peer problems (``Rather solitary, tends to play alone''), conduct problems (``Often fights with other children or bullies them''), and prosociality (``Is kind to younger children''). Greatest lower bound estimates of reliability were all greater than .95 and parallel analysis suggested that a single component was sufficient for each of the parent and teacher and age 4 and age 8 social skills factors. On the basis of this evidence, we used the total scores for these scales as developed by the LSAC administrators. Scores ranged from 0 to 10. These scores were heavily left censored for peer and conduct problems with a preponderance of students being scored as a zero by their parent or teacher. Prosocial behaviors were heavily right censored. Censoring can be viewed as a special type of missing data where scores on \(y*\) below or above the bounds are curtailed to fit within the bounds resulting in the observed scores \(y\) (Gelman, Hill, \& Vehtari, 2020). Models accounting for this censoring were used in all cases with results provided on the scale of the latent variable \(y*\) believed to underlie the censored variable (hence negative predicted scores were possible).

Parents and teachers had moderate agreement for children at age 8 with correlations ranging from \emph{r} = 0.41 95\% CI{[}0.38, 0.43{]} for peer and \emph{r} = 0.4 95\% CI{[}0.37, 0.42{]} for conduct problems to \emph{r} = 0.29 95\% CI{[}0.26, 0.31{]} for prosocial behavior.

@@

Numeracy for the child and for their complete school grade was taken from administration records of National Assessment Program -- Literacy and Numeracy (NAPLAN) test results. The relevant NAPLAN tests for this study are given to all eligible children in the country in Year 3 (age 8). The tests are scaled so they are comparable across age cohort and across year grade. These high-stakes tests have a mean of 500 and a standard deviation of 100. Included in the analysis was the first principal component of the numeracy, grammar, spelling, reading, and writing scores from NAPLAN\footnote{Parallel analysis suggested a single principal component was sufficient to account for the variation in the NAPLAN scores.}. For verbal ability (as a proxy for cognitive school readiness) we used the Peabody Picture Vocabulary test (\emph{Peabody picture vocabulary test--fourth edition}, 2007) given to participants at age 4 (i.e., at or just before school commencement). Cohort was included as a covariate in all models as was measures of rural status and gender---all measured at age 4. The sector (government or non-government) of the school the child attended was also included as a covariate.

@@

There was a relatively small amount of missing data for most variables. However, given the data were longitudinal and included both linked administrative records and data collected from both the child's teacher and parent, missing data was inevitable. Attrition was offset by using applying combined sample and attrition weights constructed for the age 8 sample. Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14\%. All other variables had 6\% missing data or less (see code book in supplementary materials ). To account for this, we constructed five imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package (Honaker, King, \& Blackwell, 2011). These imputations were used in all analyses. All analyses were done using Bayes via the BRMS package in R (Bürkner, 2017). The Bayes models were run once for each imputation and then the resulting posterior samples were pooled before point and uncertainty estimates were extracted. Interactions and their associated uncertainties were plotted by taking marginal effect estimates from a random 500 posterior draws for disadvantages (two standard deviations below the mean of SES), average, and advantaged (two standard deviations above the mean of SES) students (Gelman et al., 2020).

All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobit regression. Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails ((see Kleiber \& Zeileis, 2008). As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models. The multilevel nature of the data (children nested within schools) was addressed by the inclusion of a random intercept for schools. We clearly demarcate the units of analysis in tables and figures. I general the SDQ measures are retained on the original scale and all other continuous variables are z-scored. We retain the SDQ scores on the original scale because of the presence of norms {[}@youth2016scoring{]}. Peer and conduct problems scores of 3 or greater and prosocial scores of 7 or lower place children beyond the `close to average' group and may thus be a potential concern (Mind, 2016).

@@

In order to provide a context for the main analyses we first aimed to show: a) that children from advantaged backgrounds enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools. Student SES is slightly correlated with prosocial behaviour (0.05 95\% CI{[}0.023, 0.078{]}) but more strongly and negatively with conduct (-0.18 95\% CI{[}-0.2, -0.15{]}) and peer (-0.16 95\% CI{[}-0.19, -0.13{]}) problems. Children from lower SES backgrounds also tended to enter schools with children from similar backgrounds (0.48 95\% CI{[}0.46, 0.51{]})

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{llrrr}
\caption*{
\large Table 1\\ 
\small Effect of School Average SES on Social Skills\\ 
} \\ 
\toprule
Outcome & Report Source & Estimate & -95\% CI & +95\% CI \\ 
\midrule
conduct & teacher & $-0.152$ & $-0.261$ & $-0.046$ \\ 
social & teacher & $0.147$ & $0.056$ & $0.241$ \\ 
peer & teacher & $-0.115$ & $-0.204$ & $-0.024$ \\ 
conduct & parent & $-0.107$ & $-0.169$ & $-0.045$ \\ 
social & parent & $0.087$ & $0.018$ & $0.156$ \\ 
peer & parent & $-0.135$ & $-0.205$ & $-0.063$ \\ 
\bottomrule
\end{longtable}

@@

We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4. Results for school average SES are presented under in Table \ref{tab:table-1} (the full model results are in supplementary materials). Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (at or just prior to school enrollment) to age 8 (Year 3). School average SES negatively predicted conduct problems and positively predicted prosocial behavior for both parent and teacher reported social skills. School average SES also negatively predicted peer problems but only for parent reports.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{lllrrr}
\caption*{
\large Table 2\\ 
\small Effect of School Average SES by Student SES on Social Skills\\ 
} \\ 
\toprule
Outcome & Report Source & term & Estimate & -95\% CI & +95\% CI \\ 
\midrule
conduct & teacher & School Average SES & $-0.160$ & $-0.270$ & $-0.053$ \\ 
conduct & teacher & School Average SES by Student SES & $0.151$ & $0.064$ & $0.240$ \\ 
social & teacher & School Average SES & $0.156$ & $0.059$ & $0.252$ \\ 
social & teacher & School Average SES by Student SES & $-0.094$ & $-0.172$ & $-0.017$ \\ 
peer & teacher & School Average SES & $-0.126$ & $-0.216$ & $-0.036$ \\ 
peer & teacher & School Average SES by Student SES & $0.117$ & $0.044$ & $0.191$ \\ 
conduct & parent & School Average SES & $-0.112$ & $-0.175$ & $-0.048$ \\ 
conduct & parent & School Average SES by Student SES & $0.058$ & $0.006$ & $0.111$ \\ 
social & parent & School Average SES & $0.090$ & $0.020$ & $0.160$ \\ 
social & parent & School Average SES by Student SES & $-0.031$ & $-0.087$ & $0.027$ \\ 
peer & parent & School Average SES & $-0.144$ & $-0.217$ & $-0.074$ \\ 
peer & parent & School Average SES by Student SES & $0.117$ & $0.058$ & $0.176$ \\ 
\bottomrule
\end{longtable}

@@

The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table \ref{tab:table-2}). Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior. Interactions are plotted in Figures \ref{fig:conduct}-\ref{fig:social}. Full model results can be found in supplementary materials. The interaction plots suggest that school context effects were particularly potent for disadvantaged children but school context had minimal impact on social skills for children from advantaged backgrounds.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/peer-1.pdf}
\caption{\label{fig:peer}Interaction Plot for Peer Problems}
\end{figure}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/conduct-1.pdf}
\caption{\label{fig:conduct}Interaction Plot for Conduct Problems}
\end{figure}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/social-1.pdf}
\caption{\label{fig:social}Interaction Plot for Prosociality}
\end{figure}

@@

@@

\leavevmode\hypertarget{ref-gelman2020regression}{}%
Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press.

@@

\leavevmode\hypertarget{ref-youth2016scoring}{}%
Mind, Y. in. (2016). \emph{Scoring the strengths \& difficulties questionnaire for age 4--17 or 18+}. Retrieved from \url{https://www.ehcap.co.uk/content/sites/ehcap/uploads/NewsDocuments/236/SDQEnglishUK4-17scoring-1.PDF}

@@
